Formal Verification of Real-Time Data Processing of the LHC Beam Loss Monitoring System: A Case Study

### 2011-03-29 fmics/paper.tex
add link to proof script online
proof sketch for theorem 3

\begin{proof}
The shift register of slice $n$ is updated every $\mathsf{delay}\;n$ time steps.
When updated, the values in its cells are shifted one cell.
%The shift register cells get their values from the previous cell but they only update every $\mathsf{delay}\;n$ time steps. Therefore
Therefore\footnote{For presentation convenience, here, we do not provide the details regarding the case when $t$ is small, for example $t<m\times\mathsf{delay}\;n$.
For the full description of the theorems and proof script, please refer to {\bf Can we put a link pointing to the proof script here?}},
\begin{equation*}
\mathsf{SR}\;D\;n\;(m+1)\;t=\mathsf{SR}\;D\;n\;m\;(t-\mathsf{delay}\;n)
\end{equation*}
By induction on $m$, using the above,
\begin{equation*}
\mathsf{SR}\;D\;n\;m\;t=\mathsf{SR}\;D\;n\;0\;(t-(m\times\mathsf{delay}\;n))
\end{equation*}
By induction on $t$, we can show that $\mathsf{output}\;D\;n\;x\;t$ is a sum of values of consecutive shift register cells
\begin{equation*}
\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{\mathsf{tap}\;n\;x}\mathsf{SR}\;D\;n\;m\;t
\end{equation*}
Combining the last two results, $\mathsf{output}\;D\;n\;x\;t$ is a sum of consecutive values of the first shift register cell, $\mathsf{SR}\;D\;n\;0$.
Thus, we can express $\mathsf{output}\;D\;n\;x\;t$ in terms of $\mathsf{output}\;D\;n'\;x'\;t$, where $(n',x')=\mathsf{input}\;n$, since the source of $\mathsf{SR}\;D\;n\;0$ is the previous (input) slice's output.
Finally, by induction on $n$, we can express $\mathsf{output}\;D\;n\;x\;t$ as a function of $D$ alone, as required, using the result above for the inductive case.
\qed
\end{proof}

@@

Under such a constraint, we have proved a preliminary result bounding the error of $\mathsf{output}\;D\;n\;x\;t$ compared to $\mathsf{exact}\;D\;n\;x\;t$.
\begin{theorem}\label{max_error}
For all $k$ and $D$ satisfying $|D\;(t'+1)-D\;t'|\leq k$ at all times, the following holds for all slices $n>0$, tap positions $x$, and times $t$:
\begin{multline*}
t>(\mathsf{tap}\;n\;x+1)\times(\mathsf{delay}\;n)+\mathsf{delay}\;n+\mathsf{delay\_sum}\;n\implies\\
|\mathsf{output}\;D\;n\;x\;t-\mathsf{exact}\;D\;n\;x\;t|\leq(\mathsf{tap}\;n\;x+1)\times(\mathsf{delay\;n})\times k\times(\mathsf{delay}\;n+\mathsf{delay\_sum}\;n)
\end{multline*}
\end{theorem}
\begin{proof}
\end{proof}

>>>>

\begin{proof}
The shift register of slice $n$ is updated every $\mathsf{delay}\;n$ time steps.
When updated, the values in its cells are shifted one cell.
%The shift register cells get their values from the previous cell but they only update every $\mathsf{delay}\;n$ time steps. Therefore
Therefore\footnote{For presentation convenience, here, we do not provide the details regarding the case when $t$ is small, for example $t<m\times\mathsf{delay}\;n$.
For complete theorem statements and the proof script, please refer to \url{https://github.com/xrchz/CERN-LHC-BLMTC-SRS/blob/master/hol/srsScript.sml}.},
\begin{equation*}
\mathsf{SR}\;D\;n\;(m+1)\;t=\mathsf{SR}\;D\;n\;m\;(t-\mathsf{delay}\;n)
\end{equation*}
By induction on $m$, using the above,
\begin{equation*}
\mathsf{SR}\;D\;n\;m\;t=\mathsf{SR}\;D\;n\;0\;(t-(m\times\mathsf{delay}\;n))
\end{equation*}
By induction on $t$, we can show that $\mathsf{output}\;D\;n\;x\;t$ is a sum of values of consecutive shift register cells
\begin{equation*}
\mathsf{output}\;D\;n\;x\;t=\sum_{m=0}^{\mathsf{tap}\;n\;x}\mathsf{SR}\;D\;n\;m\;t
\end{equation*}
Combining the last two results, $\mathsf{output}\;D\;n\;x\;t$ is a sum of consecutive values of the first shift register cell, $\mathsf{SR}\;D\;n\;0$.
Thus, we can express $\mathsf{output}\;D\;n\;x\;t$ in terms of $\mathsf{output}\;D\;n'\;x'\;t$, where $(n',x')=\mathsf{input}\;n$, since the source of $\mathsf{SR}\;D\;n\;0$ is the previous (input) slice's output.
Finally, by induction on $n$, we can express $\mathsf{output}\;D\;n\;x\;t$ as a function of $D$ alone, as required, using the result above for the inductive case.
\qed
\end{proof}

@@

Under such a constraint, we have proved a preliminary result bounding the error of $\mathsf{output}\;D\;n\;x\;t$ compared to $\mathsf{exact}\;D\;n\;x\;t$.
\begin{theorem}\label{max_error}
For all $k$ and $D$ satisfying $|D\;(t'+1)-D\;t'|\leq k$ at all times, the following holds for all slices $n>0$, tap positions $x$, and times $t$:
\begin{multline*}
t>(\mathsf{tap}\;n\;x+1)\times(\mathsf{delay}\;n)+\mathsf{delay}\;n+\mathsf{delay\_sum}\;n\implies\\
|\mathsf{output}\;D\;n\;x\;t-\mathsf{exact}\;D\;n\;x\;t|\leq(\mathsf{tap}\;n\;x+1)\times(\mathsf{delay\;n})\times k\times(\mathsf{delay}\;n+\mathsf{delay\_sum}\;n)
\end{multline*}
\end{theorem}
\begin{proof}
By applying Theorem~\ref{output_eq_exact}, we are left with an equation involving $\mathsf{exact}$ but not $\mathsf{output}$, specifically between $\mathsf{exact}\;D\;n\;x\;t$ and $\mathsf{exact}\;D\;n\;x\;(\mathsf{last\_update}\;n\;t+1-\mathsf{delay\_sum}\;n)$.
Our assumption on $t$ ensures that the case where the function is equal to $0$ in Theorem~\ref{output_eq_exact} or in the definition of $\mathsf{exact}$ never applies.
Now, $\mathsf{exact}$ is defined as a sum of consecutive values of $D$, so we simply need to bound the difference between one such sum and another earlier one of the same size.
If at each step the maximum difference is $k$, then the total difference is at most $k$ times the distance between the ends of the two sums, which is the bound stated in the theorem <actually the theorem may have an additional factor increasing the size of the bound, that made it easier to prove - not sure>.
\end{proof}
